---
title: "Seabird Analysis"
author: "Lauren Davidson"
date: "14/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(here)

seabird_data <- read_csv(here("clean_data/seabird_data_clean"))
```

# Which bird had the most individual sightings?

## Interpreted as: Which bird had the highest occurence of sightings?
The Wandering Albatross, or Diomedea antipodensis, with the species abbreviation DIOANTEXU has the highest occurence of sightings, being recorded on 11,293 occasions.


```{r}
seabird_data %>% 
  select(species_abbreviation, common_name, scientific_name) %>% 
  group_by(species_abbreviation) %>% 
# Create a column for number of sightings
  mutate(num_sightings = n()) %>% 
# Ungruoup so we can see full bird name and arrange with the greatest number of sightings first.
  arrange(desc(num_sightings)) %>% 
  ungroup()
```

# Which bird had the highest total count?

The short-tailed shearwater, or Puffinus tenuirostris, abbreviated to PUFTEN, had the greatest total number of sightings at 982,553 birds spotted, due to regularly being spotted in groups of thousands of birds.

```{r}
seabird_data %>% 
  select(species_abbreviation, common_name, scientific_name, count) %>% 
  group_by(species_abbreviation) %>% 
# Create a columns totalling the number of birds seen in all sightings
  mutate(num_seen = sum(count)) %>% 
  arrange(desc(num_seen), desc(count)) %>% 
  ungroup()
```

# Which bird had the highest total count above a latitude of -30?
Further North than a latitude of -30 the bird seen the most was the wedge-tailed shearwater, or Puffinus Pacificus, seen 855 times in total.
There are only 105 recorded sightings of birds further North then -30.

```{r}
seabird_data %>% 
  select(species_abbreviation, common_name, scientific_name, count, latitude) %>%
# Only return results from -29 or greater
  filter(latitude  > -30) %>% 
  group_by(species_abbreviation) %>% 
# Create a columns totalling the number of birds seen in all sightings
  mutate(num_seen = sum(count)) %>% 
  arrange(desc(num_seen), latitude)

```

# How many different types of birds were only ever seen in groups of 1?

24 birds were only ever seen in groups of one, out of the 155 species of birds recorded.

```{r}
seabird_data %>% 
  group_by(species_abbreviation) %>% 
# Create an average of count for each bird type (as our lowest count is 1, a bird only ever seen alone will have a mean count of 1, any bird ever seen in a group will have a mean count greater than 1)
  mutate(count_avg = mean(count)) %>%
  filter(count_avg == 1) %>% 
  group_by(species_abbreviation) %>% 
  summarise()
```
```{r}
# Finding number of bird species in dataset - 155 
seabird_data %>% 
  distinct(common_name)
```

# How many penguins were seen? (Hint: there are many types of penguin)
A total of 158 penguin sightings were recorded.

```{r}
seabird_data %>% 
  select(species_abbreviation, common_name, scientific_name, count) %>%
# Only return results with the word penguin in the common name
  filter(grepl("[pP]enguin", common_name)) %>%
# Total all the numbers in the count column, i.e. the total number of penguins recorded.
  summarise(sum(count))
  
```


