---
title: "candy_analysis"
author: "Lauren Davidson"
date: "15/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(here)

candy <- read_csv(here("clean_data/clean_candy_2015_to_2017.csv"))
```

# What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
808,312 ratings of Hallowe'en treats across the three years. Treats reviewed include non-candy snacks and gifts

```{r}
# Count number of values as I have dropped non-treat questions and NA reaction
candy %>% 
  summarise(num_candy_ratings = n())
```

# What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

Those going out trick or treating have an average age of 35 years.
Those staying in have an average age of 39 years.
The average age for the dataset is 38, which is not exactly the age we expect of trick and treaters and candy eaters.

```{r}
candy %>% 
#  group_by(treating) %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```


# For each of joy, despair and meh, which candy bar revived the most of these ratings?



# How many people rated Starburst as despair?

# For the next three questions, count despair as -1, joy as +1 and meh as 0.

# What was the most popular candy bar by this rating system for each gender in the dataset?

# What was the most popular candy bar in each year?

# What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?