---
title: "candy_analysis"
author: "Lauren Davidson"
date: "15/12/2020"
output: html_document
---

```{r}
library(tidyverse)
library(here)

candy <- read_csv(here("clean_data/clean_candy_2015_to_2017.csv"))
```

# Q1: What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
772354 = number of ratings in this dataset

772354 ratings of Hallowe'en treats across the three years. Treats reviewed include non-candy snacks and gifts

```{r}
# Count number of values as I have dropped non-treat questions and NA reaction
candy %>% 
  summarise(num_candy_ratings = n())
```

# Q2: What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?
35 = average age of those going trick or treating
39 = average age of those staying in

Those going out trick or treating have an average age of 35 years.
Those staying in have an average age of 39 years.
The average age for the dataset is 38, which is not exactly the age we expect of trick and treaters and candy eaters.

```{r}
candy %>% 
  group_by(treating) %>% 
  summarise(average_age = mean(age, na.rm = TRUE))

candy %>% 
  summarise(average_age = mean(age, na.rm = TRUE))
```


# Q3: For each of joy, despair and meh, which candy bar received the most of these ratings?
Most often rated joy: 'any full size candy bar'
Most often rated despair: 'broken glow stick'
Most often rated meh: 'lollipops'

Our respondents are not very discriminating; they rated 'any full size candy bar' as joy the most often, 7589 times.
A 'broken glow stick' receieved the most despair reactions, 7905 times.
'Lollipops' are the least evocative item, receiving 1570 meh ratings.

```{r}
candy %>%
  group_by(reaction, candy) %>%
  summarise(rating_count = n()) %>%
  group_by(reaction) %>%
  slice_max(rating_count)
```


# Q4: How many people rated Starburst as despair?

1990 people rated Starburst with a despair reaction.

```{r}
candy %>% 
  filter(candy == "Starburst" &
         reaction == "DESPAIR") %>% 
  summarise(num_rating_starburst_as_despair = n())
```

# Q5: For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
rated_candy <- candy %>% 
  mutate(
    numeric_rating = case_when(
      reaction == "JOY" ~ 1,
      reaction == "DESPAIR" ~ -1,
      reaction == "MEH" ~ 0
    )
  )
```


# Q6: What was the most popular candy bar by this rating system for each gender in the dataset?
Female: 'Any full-sized candy bar'
Male: 'Any full-sized candy bar'
Other: 'Any full-sized candy bar'
NA: 'Any full-sized candy bar'

```{r}
rated_candy %>% 
  group_by(gender, candy) %>% 
  # Get a rating for each candy bar by each gender, name this "score"
  summarise(score = sum(numeric_rating), avg_rating = mean(numeric_rating)) %>% 
  # Group by gender to get a table containing all genders as rows
  group_by(gender) %>% 
  # Slice to return only the highest result
  slice_max(score)
```

# Q7: What was the most popular candy bar in each year?
2015: 'Any full-sized candy bar'
2016: 'Any full-sized candy bar'
2017: 'Any full-sized candy bar'

```{r}
rated_candy %>% 
  group_by(candy, year) %>% 
  summarise(score = sum(numeric_rating), avg_rating = mean(numeric_rating)) %>% 
  group_by(year) %>% 
  slice_max(score)
```


# Q8: What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
Most popular US: 'Any full-sized candy bar'
MOst popular Canada: 'Any full-sized candy bar'
Most popular UK: 'Cash, or other forms of legal tender'
Most popular all other countries: 'Any full-sized candy bar'


```{r}
rated_candy %>% 
  # Mutate all countries except USA, Canada and UK to NA (best way I can find to group other countries and NA together)
  mutate(country = 
          ifelse(country == "USA" | country == "UK" | country == "Canada",
          country,
          NA
          )
  ) %>% 
  # find the score for each candy by country
  group_by(country, candy) %>% 
  summarise(score = sum(numeric_rating), avg_rating = mean(numeric_rating)) %>% 
  # Group by country to return a table showing rank by country
  group_by(country) %>% 
  slice_max(score)
```

# Maybe UK people prefer cash because they're older?
Average age USA: 42
Average age UK: 38
Average age Canada: 39

Apparently not, we just like cash!

```{r}
candy %>% 
  select(country, age) %>% 
  # Mutate all countries except USA, Canada and UK to NA (best way I can find to group other countries and NA together)
  mutate(country = 
          ifelse(country == "USA" | country == "UK" | country == "Canada",
          country,
          NA
          )
  ) %>% 
  group_by(country) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```

